Layered Formal Verification of a TCP/IP Stack

### 2021-06-02 SecDev21/paper.tex
rewording

\title{Layered Formal Verification of an Industrial TCP/IP Stack}

@@

Existing research has dealt with formally verifying several protocols of TCP/IP stack levels. Early work formalized the TCP protocol to prove some key properties of the protocol itself~\cite{smith1996formal}. More recently, researchers have formally specified TCP, UDP and the Sockets API in HOL~\cite{tcp2005sigcomm,ridge2008rigorous} and used the resulting specification for testing existing libraries. Others have formally specified and implemented the SSL/TLS protocol in F*\cite{bhargavan2013implementing}, and integrated the resulting implementation in existing libraries. To the best of our knowledge, no prior work attempted to formally verify an implementation of TCP, although the authors of~\cite{ridge2008rigorous} suggest that it would possible to extract an actual implementation in Haskell from their specification.

In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in ANSI C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK~\cite{mccormick_chapin_2015}, a subset of the Ada programming language supported by formal verification tools. We chose CycloneTCP, a professional-grade open source embedded TCP/IP library developed by the Oryx Embedded company~\cite{CycloneTCP}. \Cref{Fig:TcpStack} shows the protocols supported by the library. This work focuses on the TCP and Socket components of the stack. The library's TCP implementation is meant to conform with the RFC 793 protocol specifications~\cite{rfc793}. The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report~\cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack.

@@

We present the TCP protocol in section~\ref{sec:TCP} and then briefly the
chosen stack, the Oryx Embedded one, in section~\ref{sec:stack}
before presenting the specification techniques used in section~\ref{sec:spec}
and then in sections~\ref{sec:verif} and~\ref{sec:API} we explain how we
formally implemented in SPARK the TCP user functions and how we hardened the
user's API. Finally we end in section~\ref{sec:results} by showing the results
of our work.

@@

Generally, every TCP communication session will go through the following three phases (if no error occurs): 1)~opening the connection, 2)~sending and receiving the data, and 3)~closing the connection. The behavior of the three-phase communication can be described by the state machine given in \Cref{Fig:statemachine}. An edge represents a state transition.
An edge-label in the format of $\frac{\ \ x \ \ }{\ \ y \ \ }$ represents the actions associated to the state transition, where $x$ is triggering the transition, and $y$ is produced by the transition.
The trigger $x$ is either an explicit action like \textit{Close}, or the arrival of specific flags, while $y$ always represents the flags transmitted in response to the action that caused the state transition, when present.
An explicit action is either a user-triggered action (through the user's API) or an action automatically triggered by a timeout event. The flags are embedded in the header section of a transmitted segment and can be one of the following:
\begin{itemize}
\item \ack: the last message received by the sender is acknowledged.
\item \syn: this flag is sent to establish a connection.
\item \fin: this flag is sent to close the connection.
\item \rst: this flag is sent to reset the connection, often when an error
occurred on one or the other side.
\end{itemize}

The state machine in \Cref{Fig:statemachine} does not represent the complete protocol specification; for example, it does not reflect error conditions or any actions which are not connected with the state changes. Instead, it gives an overview of all the possible states a TCP connection could reach over its lifetime. Each state has a meaning. \slisten represents waiting for a connection request from any remote TCP and port. \ssynsent and \ssynrcv represent the initialization of the connection that ends when \sestab is reached. \sfwone, \sfwtwo, \sclosing, and \stimewait are a group of states that represent waiting for a connection termination request from the remote TCP. \sclosew and \slastack represent waiting for a connection termination request from the local user. Finally, \sclosed represents no connection state at all.

@@

TCP is based on a multitasking model. Different tasks can interact and update the socket data structure, used to retain the status of a TCP connection, to handle the various events that can occur within a TCP session. RFC 793, under Section ``3.9. Event Processing'' page 52, describes a possible implementation of how to handle these events based on three tasks: one for the \emph{user calls}, one for the \emph{arriving segments} and one for the \emph{timers}. The role of each of the three tasks can be summarized as follows. User calls refer to functions, namely \texttt{open}, \texttt{close}, \texttt{abort}, \texttt{send} and \texttt{receive}, that can be called by the user to control the connection, send, or receive data. These functions can trigger transitions between the connection's states since they are intended to control the connection. In the arriving segments task, the received segments are being processed,
and the corresponding messages are sent back. Transitions between states can be triggered on the reception of a segment. Finally, timers control the timeouts, for example, the retransmission timeout to retransmit a message or the time-wait timeout to close the connection after a specified amount of time elapsed. Thus, corresponding transitions to the timeout events can also be triggered by this task. A complete description of the protocol's specification can be found in the RFC 793 document~\cite{rfc793}.

@@

\emph{GNATprove} \cite{GNATProve:users_manual} is the tool that provides formal verification for SPARK through two kinds of analyses. The first enables flow analysis to check the initialization of variables, look at data dependencies between inputs and outputs of subprograms, and detects unused assignments and unmodified variables. The second uses deductive verification to generate verification conditions for SMT solvers via a weakest-precondition calculus to detect possible runtime errors and violations of functional contracts.

@@

The verification of the entire functional specification of the \emph{CycloneTCP} library is beyond the scope of this work. Instead, the aim is to provide significant security hardening to mostly-C libraries by selective use of SPARK. Thus, this work focuses on the hardening of the TCP library areas that its original authors designated as the most vulnerable or crucial to conform to their functional specifications, namely, the Application Programming Interface (API). This mainly falls under two categories: a) proving the conformance to the protocol's functional specification for the translated code, discussed in \Cref{sec:verif}, and b) hardening the user's API to enforce its correct usage by users, discussed in \Cref{sec:API}.

Ada and SPARK offer an easy-to-use mechanism for interfacing with C code \cite{}. The mechanism allows C-written function to be called by Ada/SPARK code and vice versa and for representing shared data objects. The memory representations of a shared data object must be the same in both the C and the SPARK code. For instance, if a C function is called with a 32-bits integer argument, the SPARK interface of this function should also be called with a 32-bits integer argument. It is the responsibility of the user to ensure this.

@@

In the CycloneTCP library, pointers are used both as the type of variables to pass objects when calling other functions and as the type of structure fields; for instance, in the socket structure to store other complex data structures and recursive pointer-based data structures. SPARK uses a pointer ownership model to prevent bugs involving pointers such as memory leak or double free, which imposes some restrictions on how pointers are being manipulated. When a memory area is allocated, a pointer points to this memory area, and thus, it is considered "owned.'' The ownership of the memory can be moved, but the memory must be deallocated before the end of the execution of the program.

@@

The socket structure is shared between different functions and contains a large number of fields. Most of the fields are used in library parts that are not considered by this work, and thus, are not relevant for our verification. Thus we use \emph{ghost code} in SPARK to focus the verification effort on the relevant subset of the fields, which also makes the specifications easier to write and understand. This is a solution to the classical problem of specifying the \emph{frame condition} for a function, that is, the set of possibly modified objects. \emph{Ghost code} is used by GNATprove for analysis but is not being compiled in the final executable. More specifically, we define a \emph{model} of the actual socket as a ghost function that can be used in contracts: \spark{Model(Sock)} extracts the fields of interest from socket \spark{Sock}. A frequent usage in postconditions is

@@

This states that the relevant fields of \spark{Sock} selected by \spark{Model} are not changed by the function call, except for the field \spark{State} that is equal to \spark{TCP_STATE_CLOSED} after the call.

@@

\begin{itemize}
\item The transitions between states must respect the order given by the
state machine described in figure~\ref{Fig:statemachine}. In particular, a transition from a state to another is valid only if the state machine allows for a transition between these two states.
\item RFC 793 describes the conditions for the state of the socket which need to hold when a user's function is called~\cite[p. 52]{rfc793}. In particular, these conditions specify in what state the function can be called without returning an error. These conditions were turned into preconditions in our formalization.
\end{itemize}

@@

The original \emph{CycloneTCP} library provides no guarantees that the implementation respects the TCP functional specification. Thus, there is no assurance that the implementation will only allow valid transitions between the different states that a TCP session can exhibit. We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \Cref{Fig:statemachine}. The work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions. To achieve this, the rest of the TCP protocol, not related to the user task, also has to be taken into account. The reason for this is two-fold. First, although user-functions can trigger state transitions, the actual transitions are done by library functions that do not belong to the user-task set of functions. Second, other parts of the library can trigger state transitions during and between the user-task function calls, which can affect the intermediate and final states that a user-related function can exhibit. The TCP user API-related functions were fully translated to SPARK. SPARK bindings were also provided to most of the rest of the library's C functions to allow their invocation from the SPARK code.

RFC 793 document provides all the information needed for the allowed transitions between the different TCP possible states. For any transitions directly triggered by the call of the user-task functions, SPARK contracts, representing transitions allowed by the functional specification, are embedded within the SPARK translated code. The \texttt{TCP\_Change\_State} is a helper function that is called every time a state transition needs to be made to update the socket's state. An incorrect transition allowed by the code will be detected by the prover.

@@

The concurrent implementation of the TCP protocol allows for multiple interactions between the tasks, synchronous (with a call to \spark{TCP_Wait_For_Events}) or asynchronous (between two user function calls), and thus, numerous possible state changes at any given moment. This makes the functional specification verification significantly challenging, as SPARK does not have a native mode to deal precisely with interactions related to concurrency. To address this, we introduced SPARK functions with contracts that model how the different possible interactions can modify the state of a socket. Finding a correct and precise contract for \spark{TCP_Wait_For_Events} is challenging but needed for the verification of the TCP user functions. The solution is elaborated in the following section.

@@

% Yannick: below "a new segement...if it contains a new segment" is not phrased correctly

In the synchronous scenario of interaction between the TCP tasks, the resulting state of a user function can be affected by the rest of the TCP tasks when the user's functions release the mutex to allow for other events to occur. In this case, the function \spark{TCP_Wait_For_Events} (located in the file \texttt{tcp\_misc\_binding.ads}) is called from the user function to check if the event requested is completed. For example, the \spark{TCP_Connect} user function calls \spark{TCP_Wait_For_Events} to wait for the completion of the socket connection event, namely the \spark{SOCKET\_EVENT\_CONNECTED}. The call to the \spark{TCP_Wait_For_Events} releases the mutex, and when the \emph{arriving} task receives the appropriate segments, in this case, a segment with the \syn flag, the connection will be established. This will move the socket to the \sestab state. In the meantime, the \spark{TCP\_Update\_Events} function is monitoring for any state changes, and when it notices that the connection is established, it will trigger the \spark{SOCKET\_EVENT\_CONNECTED} event, using the OS event mechanism. This will allow the \spark{TCP\_Connect} user function to resume execution.

@@

\begin{algorithm}[t]
    \caption{Function to compute the possible state after the completion of a
particular event that is requested by a user function.}
\label{algo:waitForEvents}
\begin{algorithmic}[1]
\footnotesize
\Function{TCP\_Wait\_For\_Events\_Proof}{Socket, Event\_Mask}
    \State $S_{last} := \text{Socket}$
    \State $E :=$ \spark{TCP_Update_Events}($S_{last}$)
    \If{$(E\ \&\ \text{Event\_Mask}) \neq 0$}
        \State \textbf{return}  $S_{last}$
    \EndIf
    \For{$i=1$ \textbf{to} $3$}
        \State $S_{last} :=$ \spark{TCP_Process_One_Segment}($S_{last}$)
        \State $E :=$ \spark{TCP_Update_Events}($S_{last}$)
        \If{$(E\ \&\ \text{Event\_Mask}) \neq 0$}
            \State \textbf{return} $S_{last}$
        \EndIf
    \EndFor
    \State \textbf{return} $\emptyset$
\EndFunction
\end{algorithmic}
\end{algorithm}

@@

Essentially, the function \spark{TCP_Wait_For_Events} releases the mutex to allow for required events to happen. The function \spark{TCP_Update_Events} detects that the event's expected outcome is completed by monitoring the changes to the TCP states. Then it updates the specified event to be true in the socket structure, and it raises the desired event to allow the resuming of the user function that requested the event in the first place. Thus, the \spark{TCP_Update_Events} function is called when a segment is being received and a state change took place. This makes the \spark{TCP_Update_Events} function a perfect candidate to introduce contracts in SPARK and model the possible states after the completion of each event.

To model any possible state changes upon the reception of a segment in SPARK, the function \spark{TCP_Process_One_Segment} located in the file \texttt{tcp\_fsm\_binding.ads} is introduced.
The behavior of \spark{TCP_Wait_For_Events} is modeled by the ghost function \spark{TCP_Wait_For_Events_Proof} (that can be found in the file \texttt{tcp\_misc\_binding.ads}) shown in \Cref{algo:waitForEvents}. It computes the set of possible final states after the completion of an event by reusing the function \spark{TCP_Process_One_Segment}. Both functions have the same behavior, and thus, the contract of the function \spark{TCP_Wait_For_Events_Proof} is also attached to \spark{TCP_Wait_For_Events}. This ghost function is proved automatically without even a loop invariant, thanks to loop unrolling in GNATprove.

>>>>

\title{Layered Formal Verification of a TCP/IP Stack}

@@

Existing research has dealt with formally verifying several protocols of the TCP/IP stack. Early work formalized and proved some key properties of the TCP protocol~\cite{smith1996formal}. More recently, researchers have formally specified TCP, UDP and the Sockets API in HOL and tested existing libraries against the formal specification~\cite{tcp2005sigcomm,ridge2008rigorous}. Others have formally specified and implemented the SSL/TLS protocol in F* and integrated the resulting implementation in existing libraries~\cite{bhargavan2013implementing}. To the best of our knowledge, no prior work attempted to formally verify an implementation of TCP, although the authors of~\cite{ridge2008rigorous} suggest that it would possible to extract an actual implementation in Haskell from their specification.

In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in the C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK~\cite{mccormick_chapin_2015}, a subset of the Ada programming language supported by formal verification tools. We chose CycloneTCP, a professional-grade open source embedded TCP/IP library developed by the Oryx Embedded company~\cite{CycloneTCP}. \Cref{Fig:TcpStack} shows the protocols supported by the library. This work focuses on the TCP and Socket components of the stack. The library's TCP implementation is meant to conform to the RFC 793 protocol specifications~\cite{rfc793}. The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report~\cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack.

@@

We present the TCP protocol in section~\ref{sec:TCP} and the specification
techniques used in section~\ref{sec:spec}. We explain how we formally
implemented in SPARK the TCP user functions in section~\ref{sec:verif} and how
we hardened the user's API in section~\ref{sec:API}. Finally we conclude with a
summary of the results of our work in section~\ref{sec:results}.

@@

Generally, every TCP communication session will go through the following three phases (if no error occurs): 1)~opening the connection, 2)~sending and receiving the data, and 3)~closing the connection. The behavior of the three-phase communication can be described by the state machine given in \Cref{Fig:statemachine}. An edge represents a state transition.
An edge-label in the format of $\frac{\ \ x \ \ }{\ \ y \ \ }$ represents the actions associated with the state transition, where $x$ is triggering the transition, and $y$ is produced by the transition.
The trigger $x$ is either an explicit action like \textit{Close}, or the arrival of specific flags, while $y$ always represents the flags transmitted in response when present.
An explicit action is either a user-triggered action (through the user's API) or an action automatically triggered by a timeout event. The flags are embedded in the header section of a transmitted segment and can be one of the following:
\begin{itemize}
\item \ack: the last message received by the sender is acknowledged.
\item \syn: this flag is sent to establish a connection.
\item \fin: this flag is sent to close the connection.
\item \rst: this flag is sent to reset the connection, often when an error
occurred on one or the other side.
\end{itemize}

The state machine in \Cref{Fig:statemachine} does not represent the complete protocol specification; for example, it does not reflect error conditions or any actions which are not connected to the state changes. Instead, it gives an overview of all the possible states a TCP connection could reach over its lifetime. Each state has a clear meaning. \slisten represents waiting for a connection request from any remote TCP and port. \ssynsent and \ssynrcv represent the initialization of the connection that ends when \sestab is reached. \sfwone, \sfwtwo, \sclosing, and \stimewait are a group of states that represent waiting for a connection termination request from the remote TCP. \sclosew and \slastack represent waiting for a connection termination request from the local user. Finally, \sclosed represents no connection state at all.

@@

TCP is based on a multitasking model. Different tasks can interact and update the socket data structure, used to retain the status of a TCP connection, to handle the various events that can occur within a TCP session. RFC 793, under Section ``3.9. Event Processing'' page 52, describes a possible implementation of how to handle these events based on three tasks: one for the \emph{user calls}, one for the \emph{arriving segments} and one for the \emph{timers}. The role of each of the three tasks can be summarized as follows. User calls refer to functions, namely \texttt{open}, \texttt{close}, \texttt{abort}, \texttt{send} and \texttt{receive}, that can be called by the user to control the connection, send, or receive data. These functions can trigger transitions between the connection's states since they are intended to control the connection. In the arriving segments task, the received segments are being processed,
and the corresponding messages are sent back. Transitions between states can be triggered on the reception of a segment. Finally, timers control the timeouts, for example, the retransmission timeout to retransmit a message or the time-wait timeout to close the connection after a specified amount of time elapsed. Thus, transitions corresponding to the timeout events can also be triggered by this task. A complete description of the protocol's specification can be found in the RFC 793 document~\cite{rfc793}.

@@

\emph{GNATprove} \cite{GNATProve:users_manual} is the tool that provides formal verification for SPARK through two kinds of analyses. The first enables flow analysis to check the initialization of variables, look at data dependencies between inputs and outputs of subprograms, and detect useless code. The second uses deductive verification to generate verification conditions for SMT solvers via a weakest-precondition calculus to detect possible runtime errors and violations of functional contracts.

@@

The verification of the entire functional specification of the CycloneTCP library is beyond the scope of this work. Instead, the aim is to provide significant security hardening to mostly-C libraries by selective use of SPARK. Thus, this work focuses on the hardening of the TCP library areas that its original authors designated as the most vulnerable or crucial to conform to their functional specifications, namely, the Application Programming Interface (API). This mainly falls under two categories: a) proving the conformance to the protocol's functional specification for the translated code, discussed in \Cref{sec:verif}, and b) hardening the user's API to enforce its correct usage by users, discussed in \Cref{sec:API}.

Ada and SPARK offer an easy-to-use mechanism for interfacing with C code. The mechanism allows C-written function to be called by Ada/SPARK code and vice versa and for representing shared data objects. The memory representations of a shared data object must be the same in both the C and the SPARK code. For instance, if a C function is called with a 32-bits integer argument, the SPARK interface of this function should also be called with a 32-bits integer argument. It is the responsibility of the user to ensure this.

@@

In the CycloneTCP library, pointers are used both as the type of variables to pass objects when calling other functions and as the type of structure fields; for instance, in the socket structure to store other complex data structures and recursive pointer-based data structures. SPARK uses a pointer ownership model to prevent bugs involving pointers such as memory leak or double free, which imposes some restrictions on how pointers are being manipulated. When a memory area is allocated, a pointer points to this memory area, and thus, it is considered ``owned.'' The ownership of the memory can be moved, but the memory must be deallocated before the end of the execution of the program.

@@

The socket structure is shared between different functions and contains a large number of fields. Most of the fields are used in library parts that are not considered by this work, and thus, are not relevant for our verification. Thus we use \emph{ghost code} in SPARK to focus the verification effort on the relevant subset of the fields, which also makes the specifications easier to write and understand. This is a solution to the classical problem of specifying the \emph{frame condition} for a function, that is, the set of objects possibly modified by the function. \emph{Ghost code} is used by GNATprove for analysis but is not being compiled in the final executable. More specifically, we define a \emph{model} of the actual socket as a ghost function that can be used in contracts: \spark{Model(Sock)} extracts the fields of interest from socket \spark{Sock}. A frequent usage in postconditions is

@@

\noindent
This states that the relevant fields of \spark{Sock} selected by \spark{Model} are not changed by the function call, except for the field \spark{State} that is equal to \spark{TCP_STATE_CLOSED} after the call.

@@

\begin{itemize}
\item The transitions between states must respect the order given by the
state machine described in \Cref{Fig:statemachine}. In particular, a transition from a state to another is valid only if the state machine allows for a transition between these two states.
\item RFC 793 describes the conditions for the state of the socket which need to hold when a user's function is called~\cite[p. 52]{rfc793}, to avoid errors. These conditions were turned into preconditions in our formalization.
\end{itemize}

@@

The original CycloneTCP library provides no guarantees that the implementation respects the TCP functional specification. Thus, there is no assurance that the implementation will only allow valid transitions between the different states that a TCP session can exhibit. We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \Cref{Fig:statemachine}. Although this work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions, the rest of the TCP protocol has to be taken into account. The reason for this is two-fold. First, although TCP user functions can trigger state transitions, the actual transitions are done by other library functions called from TCP user functions. Second, other parts of the library can trigger state transitions during and between TCP user function calls, which can affect the intermediate and final states that a TCP user function can exhibit.
The TCP user functions were fully translated to SPARK. SPARK bindings were also provided for most of the rest of the library's C functions to allow their invocation from the SPARK code.

RFC 793 document provides all the information needed for the allowed transitions between the different TCP possible states. For any transitions directly triggered by calling the TCP user functions, SPARK contracts, representing transitions allowed by the functional specification, are embedded within the SPARK translated code. The helper function \texttt{TCP\_Change\_State} is called explicitly every time a state transition needs to be made to update the socket's state. Any incorrect transition allowed by the code will be detected by GNATprove.

@@

The concurrent implementation of the TCP protocol allows for multiple interactions between the tasks, both synchronous (with a call to \spark{TCP_Wait_For_Events}) or asynchronous (between two user function calls), and thus, numerous possible state changes at any given moment. This makes the functional specification verification significantly challenging, as SPARK does not have a native mode to deal precisely with interactions related to concurrency. To address this, we introduced SPARK functions with contracts that model how the different possible interactions can modify the state of a socket. Finding a correct and precise contract for \spark{TCP_Wait_For_Events} was challenging but needed for the verification of the TCP user functions. The solution is elaborated in the following section.

@@

In the synchronous scenario of interaction between the TCP tasks, the resulting state of a user function can be affected by the rest of the TCP tasks when the TCP user function releases the mutex to allow for other events to occur. In this case, the function \spark{TCP_Wait_For_Events} is called from the TCP user function to check if the event requested is completed. For example, the \spark{TCP_Connect} user function calls \spark{TCP_Wait_For_Events} to wait for the completion of the socket connection event, namely \spark{SOCKET\_EVENT\_CONNECTED}. The call to \spark{TCP_Wait_For_Events} releases the mutex, and when the relevant task receives the appropriate segment, in this case, a segment with the \syn flag, the connection will be established. This will move the socket to the \sestab state. In the meantime, the \spark{TCP\_Update\_Events} function is monitoring for any state changes, and when it notices that the connection is established, it will trigger the \spark{SOCKET\_EVENT\_CONNECTED} event, using the OS event mechanism. This will allow the \spark{TCP\_Connect} user function to resume execution.

@@

\begin{algorithm}[t]
    \caption{Function to compute the possible state after the completion of a
particular event that is requested by a TCP user function.}
\label{algo:waitForEvents}
\begin{algorithmic}[1]
\footnotesize
\Function{TCP\_Wait\_For\_Events\_Proof}{Socket, Event\_Mask}
    \State $S_{last} := \text{Socket}$
    \State $E :=$ \spark{TCP_Update_Events}($S_{last}$)
    \If{$(E\ \&\ \text{Event\_Mask}) \neq 0$}
        \State \textbf{return}  $S_{last}$
    \EndIf
    \For{$i=1$ \textbf{to} $3$}
        \State $S_{last} :=$ \spark{TCP_Process_One_Segment}($S_{last}$)
        \State $E :=$ \spark{TCP_Update_Events}($S_{last}$)
        \If{$(E\ \&\ \text{Event\_Mask}) \neq 0$}
            \State \textbf{return} $S_{last}$
        \EndIf
    \EndFor
    \State \textbf{return} $\emptyset$
\EndFunction
\end{algorithmic}
\end{algorithm}

@@

Essentially, the function \spark{TCP_Wait_For_Events} releases the mutex to allow for required events to happen. The function \spark{TCP_Update_Events} detects that the event's expected outcome is completed by monitoring the changes to the TCP states. Then it updates the specified event to be true in the socket structure, and it raises the desired event to allow resuming the TCP user function that requested the event in the first place. Thus, the \spark{TCP_Update_Events} function is called when a segment is being received and a state change took place. This makes the \spark{TCP_Update_Events} function a perfect candidate to introduce contracts in SPARK and model the possible states after the completion of each event.

To model any possible state changes upon the reception of a segment in SPARK, the function \spark{TCP_Process_One_Segment} is introduced.
The behavior of \spark{TCP_Wait_For_Events} is modeled by the ghost function \spark{TCP_Wait_For_Events_Proof} shown in \Cref{algo:waitForEvents}. It computes the set of possible final states after the completion of an event by calling repeatedly the function \spark{TCP_Process_One_Segment}. Both functions have the same behavior, and thus, the contract of the function \spark{TCP_Wait_For_Events_Proof} is also attached to \spark{TCP_Wait_For_Events}. This ghost function is proved automatically without even requiring a loop invariant, thanks to loop unrolling in GNATprove.
